# Reproduceable Research - Course Project-1
============================================
Author: Narayan

Modified date : 12/27/2017

## Hitogram of number of steps taken per day

```{r echo=FALSE}
MyData <- read.csv(file="./data/activity.csv", header=TRUE, sep=",") ## Read file
MyData$date <- as.Date(MyData$date)
totSteps <- with(MyData, tapply(steps, date, sum, na.rm = T)) 
hist(totSteps, xlim = c(0,25000), ylim=c(0,30), main="Histogram of steps taken each day", breaks=10, las=1, border="blue", col="gray", 
     xlab="Total Number of Steps taken")
```

#### Calculate Mean and Median of the total number of steps taken per day
#### print the values in an HTML table and create chart for Average number of steps taken per day.

```{r showtable, results='asis', echo=TRUE}
library(xtable)

medianWithoutNA <- function(x) {
   median(x[which(!is.na(x) & x > 0)])
}

avgStepsPerDay <- with(MyData, tapply(steps, date, mean, na.rm = T))
medStepsPerDay <- with(MyData, tapply(steps, date, medianWithoutNA))
d1 <- data.frame(date = names(avgStepsPerDay), avgSteps=avgStepsPerDay)
d1$medSteps <- medStepsPerDay
rownames(d1) <- c()
xt <- xtable(d1)
names(xt) <- c('Date', 'Mean of steps Taken per Day','Median of steps Taken per Day')
print(xt, type="html")

plot(as.Date(names(avgStepsPerDay),"%Y-%m-%d"),avgStepsPerDay, type="l", xlab="Activity Dates", ylab="Averages steps per Day") ## Time series plot
mtext("Time Series Chart - Average number of steps per day",side=3, line=2)
```

#### Calculate and print the 5 minute interval that on average contains the max number of steps
#### Step1 : calculated mean value per each 5 minute interval
#### Step 2: find the max value of the mean values and the corresponding interval value
#### Step 3: print both values

```{r echo=TRUE}
avgStepsPerInterval <- with(MyData, tapply(steps, interval, mean, na.rm = T))
d2 <- data.frame(interval = names(avgStepsPerInterval), avgSteps=avgStepsPerInterval)
maxMeanVal <- max(d2$avgSteps)
intrname <- as.character(d2[d2$avgSteps==maxMeanVal,1])
sprintf("5 minute interval that contains maximum number of steps (%f)on average: %s", maxMeanVal,intrname )
```


#### Impute values for missing values by filling in monthly mean values
#### Calculate the total number of steps taken again and create a histogram again
#### Process data by intervals and divide data into weekend and weekdays data sets. Plot time series charts

```{r echo=TRUE}
library(chron)

p1 <- MyData[!is.na(MyData$steps) & as.character(MyData$date,"%m") == "10",]
OctMean <- mean(p1$steps, na.rm=TRUE)
MyData[is.na(MyData$steps) & as.character(MyData$date,"%m") == "10",]$steps = OctMean

p2 <- MyData[!is.na(MyData$steps) & as.character(MyData$date,"%m") == "11",]
NovMean <- mean(p2$steps, na.rm=TRUE)
MyData[is.na(MyData$steps) & as.character(MyData$date,"%m") == "11",]$steps = NovMean 

totSteps <- with(MyData, tapply(steps, date, sum, na.rm = T)) 
hist(totSteps, xlim = c(0,25000), ylim=c(0,30), main="Histogram of steps taken each day (post imputing)", breaks=10, las=1, border="blue", col="gray")

WEMyData <- (MyData[is.weekend(MyData$date),])
WDMyData <- (MyData[!is.weekend(MyData$date),])
WEGraph <- tapply(WEMyData$steps,WEMyData$interval,mean)
WDGraph <- tapply(WDMyData$steps,WDMyData$interval,mean)
d2 <- data.frame(interval=names(WEGraph), avgSteps=WEGraph)

par(mfrow=c(2,1), mgp=c(.25,.25,.25), mar=c(3, 3, 1, 1))
plot(names(WEGraph), WEGraph, type="l", axes=FALSE, ann= FALSE, main="par(mar=c(1, 2, 2, 1))")
axis(2, cex.axis=.7)
box()
mtext("Number of Steps", side=2, line=1.3)
mtext("Weekend", side=3, line=0)

plot(names(WDGraph), WDGraph, type = "l", axes=FALSE, ann= FALSE, main="par(mar=c(3, 2, 1, 1))")
axis(1, cex.axis=.7)
axis(2, cex.axis=.7)
box()
mtext("Interval", side=1, line=1.3)
mtext("Number of Steps", side=2, line=1.3)
mtext("Weekday", side=3, line=0)
```